---
title: "R Notebook"
output: html_notebook
---

```{r}
library(MASS)
```


#Q2. What predictors are most important in telling how a game ends - using LDA, QDA, Multinomial - Chi-square test

```{r}
chess_data = read.csv("Q2_final_training_data.csv", header = TRUE)
chess_data
```

#dropping columns not required for lda i.e. W_first5, B_first5
```{r}
#chess_data2 = subset(chess_data, select = -c(X,P21,W_First5,B_First5,ID_unit,Stratum,Prob) )
chess_data2 = subset(chess_data, select = c(rated,turns,victory_status,winner,pace_type,Rating_WB,opening_strategy) )
chess_data2
```



```{r}
lda.fit<-lda(victory_status~., data=chess_data2)
lda.fit
#summary(lda.fit)
```
```{r}
summary(lda.fit)
```


```{r}
plot(lda.fit)
```

```{r}
# test = read.csv('test_lda_q2.csv', header = TRUE)
# test
```



```{r}
#predict with test
# lda.pred=predict(lda.fit, test)
# names(lda.pred)
```

```{r}
#table(lda.pred$class, Test$victory_status)
```


```{r}
chess_data2$victory_status = as.factor(chess_data2$victory_status)
contrasts(chess_data2$victory_status)

```

```{r}
library(klaR)
partimat(victory_status~., data = chess_data2, method="lda")
```

#cross validation
```{r}
library(caret)
```
```{r}
set.seed(5)
folds<-createFolds(chess_data2$victory_status, k=10)
```


```{r}
fold1<-chess_data2[folds$Fold1,]
table(fold1$victory_status)
```
#a function to calculate the misclassification rate (LDA).
```{r}
library (MASS)
misclassification<-function(idx){
train<-chess_data2[-idx,]
test<-chess_data2[idx,]
fit<-lda(victory_status~., data=train)
pred<-predict(fit,test,type = "class")
return(1-mean(pred$class==test$victory_status))
}
```
#apply the function “lapply” to pass along a vector of indices to the function “misclassification”
```{r}
mis_rate=lapply(folds,misclassification)
mis_rate
```
#calculate the average misclassification error.
```{r}
mean(as.numeric(mis_rate))
```


# #######################

#Q1. Which columns predict if white or black wins.

```{r}
chess_data_q1_raw = read.csv("q1_final_training_data.csv", header = TRUE)
chess_data_q1_raw
```

```{r}
chess_data_q1 = subset(chess_data_q1_raw, select = -c(X,W_First5,B_First5,ID_unit,Stratum,Prob,rating_bins) )
```


```{r}
lda.fit<-lda(winner~., data=chess_data_q1)
lda.fit
```

```{r}
plot(lda.fit)
```

```{r}
# test = read.csv('test_lda_q2.csv', header = TRUE)
# test
```


```{r}
#predict with test
# lda.pred=predict(lda.fit, test)
# names(lda.pred)
```

```{r}
#table(lda.pred$class, test$winner)
```

#cross validation
```{r}
library(caret)
```

```{r}
set.seed(5)
folds<-createFolds(chess_data_q1$winner, k=10)
```


```{r}
fold1<-chess_data_q1[folds$Fold1,]
table(fold1$winner)
```
#a function to calculate the misclassification rate (LDA).
```{r}
library (MASS)
misclassification<-function(idx){
train<-chess_data_q1[-idx,]
test<-chess_data_q1[idx,]
fit<-lda(winner~., data=train)
pred<-predict(fit,test,type = "class")
return(1-mean(pred$class==test$winner))
}
```
#apply the function “lapply” to pass along a vector of indices to the function “misclassification”
```{r}
mis_rate=lapply(folds,misclassification)
mis_rate
```
#calculate the average misclassification error.
```{r}
mean(as.numeric(mis_rate))
```


#Classification tree

```{r}
library(tree)
```

```{r}
chess_data_q1_raw = read.csv("q1_final_training_data.csv", header = TRUE)
chess_data_q1_raw
```

```{r}
chess_data_q1_tree = subset(chess_data_q1_raw, select = c(rated,turns, victory_status,pace_type, W1, B1,winner, turns, Rating_WB, WOE_Opening_Name) )

```

```{r}
str(chess_data_q1_tree)
```
```{r}
#factor everything
#chess_data_q1_tree$op_name = as.factor(chess_data_q1_tree$op_name)
chess_data_q1_tree$winner = as.factor(chess_data_q1_tree$winner)
chess_data_q1_tree$victory_status = as.factor(chess_data_q1_tree$victory_status)
chess_data_q1_tree$rated = as.factor(chess_data_q1_tree$rated)

chess_data_q1_tree$pace_type = as.factor(chess_data_q1_tree$pace_type)
chess_data_q1_tree$B1 = as.factor(chess_data_q1_tree$B1)
chess_data_q1_tree$W1 = as.factor(chess_data_q1_tree$W1)
# chess_data_q1_tree$W_First5 = as.factor(chess_data_q1_tree$W_First5)
# chess_data_q1_tree$B_First5 = as.factor(chess_data_q1_tree$B_First5)
#chess_data_q1_tree$opening_strategy = as.factor(chess_data_q1_tree$opening_strategy)
#chess_data_q1_tree$rating_bins = as.factor(chess_data_q1_tree$rating_bins)

```

```{r}
contrasts(chess_data_q1_tree$rated)

```

```{r}
# No_salary=subset(Hitters,is.na(Hitters$Salary))
# New_data=na.omit(Hitters)
```


```{r}
tree.class<-tree(factor(winner)~., data = chess_data_q1_tree)
summary(tree.class)
```

```{r}
plot(tree.class)
text(tree.class, pretty=0)
```

```{r}
# tree.pred<-predict(tree.class,test,type = "class")
# table(tree.pred,test$winner)
```

```{r}
set.seed(5)
cv.class<-cv.tree(tree.class, FUN = prune.misclass, K=10) 
plot(cv.class$size, cv.class$dev,type="b")
```
```{r}
prune.class=prune.tree(tree.class,best=2)
plot(prune.class)
text(prune.class,pretty=0)
```

#testing after pruning
```{r}
# prune.pred=predict(prune.class,test,type="class")
# table(prune.pred,test$winner)
```

#cv error classification

```{r}
set.seed(5)
folds<-createFolds(chess_data_q1_tree$winner, k=10)

library (MASS)
misclassification2<-function(idx){
train<-chess_data_q1_tree[-idx,]
test<-chess_data_q1_tree[idx,]
fit2<-tree (winner~., data=chess_data_q1_tree)
pred2<-predict (fit2, test,type = "class")
print(table(pred2,test$winner)) 
return(1-mean(pred2==test$winner))
}
```

```{r}
mis_rate2=lapply(folds,misclassification2)
mis_rate2
```
```{r}
mean(as.numeric(mis_rate2))
```
